Implications of animal movement rate variation in spatial patterns of seed dispersal

### 2021-08-18 Ch1_movement_rates/Ch1_markdown.Rmd
so many updates.

```{r message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

@@

```
```{r}
sim_movement <- function(prm, t = 1000, plot.it = TRUE, return.data.frame = FALSE){
  tru.rate <- round(prm, 3)
  movedist <- rexp(t, rate = 1/tru.rate)
  angle <- runif(t, min = 0, max = 360)
  distx <- movedist*cos(angle)
  xloc <- c(0, cumsum(distx))
  disty <- movedist*sin(angle)
  yloc <- c(0, cumsum(disty))
  animalTraj <- data.frame(time = 0:t, xloc = xloc, yloc = yloc)
  if(plot.it == TRUE){
    plot(x = xloc, y = yloc, type = "l", main = paste("Rate=",tru.rate))
  }
  if(return.data.frame == TRUE){
    return(animalTraj)
  }
}
```

@@

Population1
```{r}

@@

for(j in 1:n.individuals){
  for(k in 1:3){
  a <- sim_movement(m_1[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

Population2
```{r}

for(j in 1:individual){
  for(k in 1:10){
  a <- sim_movement(m_2[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

Population3
```{r}

for(j in 1:individual){
  for(k in 1:10){
  a <- sim_movement(m_3[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

@@

    df <- rbind.data.frame(df, a) 
    summ.df <- rbind.data.frame(summ.df, b) 
}
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, 
                group = run,
                color = indiv)) +
  # facet_wrap(~run) +
  geom_path() +
  scale_color_manual(values = mycols) +
  geom_point(data = df %>% drop_na(s.id), 
             aes(x = xloc, y = yloc)) +
  geom_point(data = summ.df, aes(x = x, y=y), color = "black") +
  theme_bw()

df %>% 
  ggplot(., aes(x = xloc, y = yloc)) +
  # stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white") +
  geom_bin2d() +
  # geom_hex() +
  scale_fill_continuous(type = "viridis") +
  theme_bw()

summ.df %>% 
  ggplot(., aes(y = dsprsn)) +
  geom_boxplot()

summ.df %>% 
  ggplot(., aes(x = av.disp)) +
  geom_density(color = "blue") + # for the average dispersal per run
  geom_density(data = df %>% drop_na(s.id),
                 aes(x = disp)) + # for every seed
  theme_bw()
```

@@

for(j in 1:individual){
  for(k in 1:nruns){
    a <- sim_seeds(m.prms = m_1[j]) %>% 
      mutate(indiv = as.factor(j),
             run = factor(paste0("r_", k), levels = paste0("r_", 1:nruns)))

@@

Visualize animal movement and seed dispersal 
```{r}
df %>% 
  ggplot(., aes(x = xloc, y = yloc)) +
  facet_wrap(~indiv) +
  # stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white") +
  geom_bin2d() +
  # geom_hex() +
  # scale_fill_continuous(type = "viridis") +
  # geom_point(data = df %>% drop_na(s.id), 
  #            aes(x = xloc, y = yloc)) +
  geom_point(data = summ.df, aes(x = x, y=y), color = "red") +
  theme_bw()
```
Seed dispersal kernels
```{r}
df %>% 
  drop_na(s.id) %>% 
  ggplot(., aes(x = disp, color = indiv)) +
  geom_density() +
  scale_color_manual(values = mycols) +
  theme_bw()

@@

Dispersion
```{r}
summ.df %>% 
  ggplot(., aes(y = dsprsn, x = indiv)) + 
  geom_boxplot()
```
```{r}
s.df <- df %>% drop_na(s.id) %>%
  group_by(indiv) %>%
  sample_n(., 50)

>>>>

```{r message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
```

@@

```{r}
sim_movement <- function(prm, t = 1000, plot.it = TRUE, return.data.frame = FALSE){
  tru.rate <- round(prm, 3)
  movedist <- rexp(t, rate = 1/tru.rate)
  angle <- runif(t, min = 0, max = 360)
  distx <- movedist*cos(angle)
  xloc <- c(0, cumsum(distx))
  disty <- movedist*sin(angle)
  yloc <- c(0, cumsum(disty))
  animalTraj <- data.frame(time = 0:t, xloc = xloc, yloc = yloc)
  if(plot.it == TRUE){
    plot(x = xloc, y = yloc, type = "l", main = paste("Rate=",tru.rate))
  }
  if(return.data.frame == TRUE){
    return(animalTraj)
  }
}
```

@@

## Movement for individuals in the three populations
```{r eval=FALSE}
nruns <- 3
# Population1
df <- NULL

for(j in 1:n.individuals){
  for(k in 1:nruns){
  a <- sim_movement(m_1[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, group = run, color = indiv)) +
  geom_path() +
  scale_color_manual(values = mycols) +
  theme_bw()

# Population2
df <- NULL

for(j in 1:n.individuals){
  for(k in 1:nruns){
  a <- sim_movement(m_2[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, group = run, color = indiv)) +
  geom_path() +
  scale_color_manual(values = mycols) +
  theme_bw()

# Population3
df <- NULL

for(j in 1:n.individuals){
  for(k in 1:nruns){
  a <- sim_movement(m_3[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste("r_", k))
  df <- rbind.data.frame(df, a)
}
}

```{r}
nruns <- 3
m.data <- data.frame(m_1, m_2, m_3)

@@

for(m in 1:3){
  m.0 <- m.data[m]
  for(j in 1:n.individuals){
  for(k in 1:nruns){
  a <- sim_movement(m_2[j], plot.it = FALSE, return.data.frame = TRUE) %>% 
    mutate(indiv = as.factor(j),
           run = paste0("r_", k),
           pop.id = paste0("p_",m))
  df <- rbind.data.frame(df, a)
}
}

}

df %>% 
  ggplot(., aes(x = xloc, y=yloc, group = run, color = indiv)) +
  facet_wrap(~pop.id) +
  geom_path() +
  scale_color_manual(values = mycols) +
  theme_bw() +
  theme(legend.position = "bottom") +
  NULL
```

@@

df %>% 
  ggplot(., aes(x = xloc, y=yloc, 
                group = run,
                color = indiv)) +
  # facet_wrap(~run) +
  geom_path() +
  scale_color_manual(values = mycols) +
  geom_point(data = df %>% drop_na(s.id), 
             aes(x = xloc, y = yloc)) +
  geom_point(data = summ.df, aes(x = x, y=y), color = "black") +
  theme_bw() +
  labs(title = "Animal seed droppings", caption = "Lines show an individual's trajectory.\n Red dots show all seed droppings.\n Black dots show average location of seed per run") +
  theme(legend.position = "none") -> p1
# p1

df %>% 
  ggplot(., aes(x = xloc, y = yloc)) +
  # stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white") +
  geom_bin2d() +
  # geom_hex() +
  # scale_fill_continuous(type = "viridis") +
  scale_fill_gradient(low = "grey", high = "black") +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(title = "Density of animal movement") -> p2
p2

cowplot::plot_grid(p1, p2)

# summ.df %>% 
#   ggplot(., aes(y = dsprsn)) +
#   geom_boxplot()
# 
# summ.df %>% 
#   ggplot(., aes(x = av.disp)) +
#   geom_density(color = "blue") + # for the average dispersal per run
#   geom_density(data = df %>% drop_na(s.id),
#                  aes(x = disp)) + # for every seed
#   theme_bw()
```

@@

for(j in 1:n.individuals){
  for(k in 1:nruns){
    a <- sim_seeds(m.prms = m_1[j]) %>% 
      mutate(indiv = as.factor(j),
             run = factor(paste0("r_", k), levels = paste0("r_", 1:nruns)))

@@

Visualize animal movement and seed dispersal. It is kind of obvious, but individuals that move further away, also disperse seeds further away and scatter them more. Red points are every seed dispersed, black points show the average seed location per run. Each panel corresponds to one individual disperser.
```{r}
df %>% 
  ggplot(., aes(x = xloc, y = yloc)) +
  facet_wrap(~indiv) +
  # stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white") +
  geom_bin2d() +
  # geom_hex() +
  # scale_fill_continuous(type = "viridis") +
  scale_fill_gradient(low = "grey", high = "black") +
  geom_point(data = df %>% drop_na(s.id),
             aes(x = xloc, y = yloc), color = "red", alpha = 0.3) +
  geom_point(data = summ.df, aes(x = x, y=y), color = "black") +
  theme_bw()
```

Seed dispersal kernels produced by each individual.
```{r}
df %>% 
  drop_na(s.id) %>% 
  ggplot(., aes(x = disp, color = indiv)) +
  geom_density() +
  scale_color_manual(values = mycols) +
  theme_bw()

@@

Dispersion. We calculate dispersion for every simulated run. Every run means an individual disperser gets 20 seeds, they move until they drop all seeds. Then, dispersion measures how far away from each other every seed is. 
```{r}
summ.df %>% 
  ggplot(., aes(y = dsprsn, x = indiv)) + 
  geom_boxplot()
```
```{r}
s.df <- df %>% drop_na(s.id) %>%
  group_by(indiv) %>%
  sample_n(., 50)
